---
title: "Unit_9_Assignment"
author: "Matt Chinchilla"
date: "7/6/2019"
output: html_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
#Libraries
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
```


#### A. Clean an prepare the Data:

##### Download Data
```{r}
#Create Data frame for Beers.csv data
Beers <- read.csv("./Beers.csv")

#Create data frame for Breweries.cvs
Breweries <- read.csv("./Breweries.csv")
```


##### 2. Merge Data into data frame
```{r}
#Merge into new dataframe BeersBreweries
BeersBreweries <- inner_join(Beers,Breweries, by = c("Brewery_id" = "Brew_ID"))

#Rename columns from Name.x and Name.y to Beer_Name and Brewery_Name 
names(BeersBreweries)[1] <- c("Beer_Name")
names(BeersBreweries)[8] <- c("Brewery_Name")
```


##### 3. Clean white spaces in State abbreviations.
```{r}
#trim the white space in state abbreviations 
BeersBreweries$State <- trimws(BeersBreweries$State)
```


##### 4. Create Data frame beerCOTX with only Colorado and Texas beers. Exclude NA's from new data frame
```{r}
#Create subset of BeersBreweries with only Texas and Colorado Breweries 
beerCOTX <- filter(BeersBreweries, State == c("TX","CO"))

#Omit all rows from beerCOTX with null value for IBU
beerCOTX <- beerCOTX %>% na.omit(beerCOTX$IBU)
```


##### 5. Order beerCOTX by IBU asscending. View data.
```{r}
#Arrange Rows in ascending order by IBU
beerCOTX <- beerCOTX %>% arrange(IBU)

#View beerCOTX
kable(head(beerCOTX)) %>% kable_styling(bootstrap_options = "striped")
```

#### B. 

##### 6. Create Initial plots of Texas and Colorado Data
```{r}
#Plot IBU and ABV for all values in beerCOTX and assign to variable p
p <- ggplot(beerCOTX, aes(IBU, ABV)) + geom_point()

#Split the previous plot into two based on state
p + facet_grid(State ~.)
```


#### C. Model the data

##### 7a. Texas model
```{r}
#Split beerCOTX by state

#Texas Breweries
beerTX <- filter(beerCOTX, State == "TX")

#Calculate coeficients
lmtx <- lm(ABV ~ IBU, data = beerTX)

#Grab coefficients data
coefs <- coef(lmtx)

#Plot using base plot function.
plot(beerTX$IBU, beerTX$ABV, pch=20, col = "red", xlab = "IBU", ylab = "ABV", main = "ABV vs IBU Texas")
abline(coefs[1], coefs[2], col = "blue")
```

##### 7b. Colorado Model
```{r}
#Split beerCOTX by state

#Colorado Breweries
beerCO <- filter(beerCOTX, State == "CO")

#Calculate coeficients
lmco <- lm(ABV ~ IBU, data = beerCO)

#Grab coefficients data
coefs1 <- coef(lmco)

#Plot using base plot function.
plot(beerCO$IBU, beerCO$ABV, pch=20, col = "red", xlab = "IBU", ylab = "ABV", main = "ABV vs IBU Colorado")
abline(coefs1[1], coefs1[2], col = "blue")
```
##### 8. Address the assumptions of the regression model.

##### 8a. 
```{r}
#Texas
length(beerTX$Beer_ID)

#Colorado
length(beerCO$Beer_ID)
```


##### 8b. The normal distributions for Texas and Colorado do not have equal standard deviations.
```{r}
# Standard deviation Texas
sd(beerTX$ABV)

# Standard deviation Colorado
sd(beerCO$ABV)
```
##### 8c.The mean of the distributions have a linear relationship to IBU??? Also not sure how to do this


##### 8d. Independance ??? Not sure how to do this


#### D. Gain inference from the model

##### 9a. The Texas model has an R^2 variance of 65% adjusted to 64%. This shows me the model can explane over 50% of the variance in ABV.
```{r}
#Parameter estimate table Texas
tab_model(lmtx)
```

##### 9b. The Colorado model has an R^2 variance of 34.5% adjusted to 33.6%. This shows me the model can explane less than 50% of the variance in ABV. This is probably not a good model for Colorado.
```{r}
#Parameter estimate table Colorado
tab_model(lmco)
```

##### 9c. Based on my observations there does seem to be evidence that the relationship between ABV and IBU are diffent in Texas and Colorado.

##### 10 I have no idea???

#### E. Compare two competing models

##### 11. Create a new column IBU2 with the square of IBU for each row. 
```{r}
#Add column IBU2 to beerCOTX data frame.
beerCOTX$IBU2 <- beerCOTX$IBU ^ 2

#Confirm new column
kable(head(beerCOTX)) %>% kable_styling(bootstrap_options = "striped")
```

##### 12a. Split the beerCOTX data into training and test data frames based on state

##### Texas training and test data frames
```{r}
# Filter on Texas data
beerTX <- filter(beerCOTX, State == "TX")

# Create random sample set of Texas training observations based on 60% data for training
TrainObs = sample(seq(1,dim(beerTX)[1]),round(.60*dim(beerTX)[1]), replace = FALSE)

# Create data frame of training observations
TrainingTX <- beerTX[TrainObs,]

# Summary of TrainingTX
summary(TrainingTX)

# Create a data frame of test observations
TestTX <- beerTX[-TrainObs,]

# Summary of TrainingTX
summary(TestTX)
```

##### Colorado training and test data frames
```{r}
# Filter on Texas data
beerCO <- filter(beerCOTX, State == "CO")

# Create random sample set of Texas training observations based on 60% data for training
TrainObs1 = sample(seq(1,dim(beerCO)[1]),round(.60*dim(beerCO)[1]), replace = FALSE)

# Create data frame of training observations
TrainingCO <- beerCO[TrainObs1,]

# Summary of TrainingCO
summary(TrainingCO)

# Create a data frame of test observations
TestCO <- beerCO[-TrainObs1,]

# Summary of TestCO
summary(TestCO)
```

##### 13 compare model1 and model2

##### Model 1 Texas
```{r}
#Train model1 Texas
model1tx <- lm(ABV ~ IBU, data = TrainingTX)

#Predictions based on model1tx
model1tx_preds <- predict(model1tx, newdata = TestTX)
as.data.frame(model1tx_preds)

#Model1tx MSPE
MSPETX1 = mean((TestTX$ABV - model1tx_preds)^2)
MSPETX1
```
##### Model 2 Texas
```{r}
#Train model2 Texas
model2tx <- lm(ABV ~ IBU + IBU2, data = TrainingTX)

#Predictions based on model1tx
model2tx_preds <- predict(model2tx, newdata = TestTX)
as.data.frame(model2tx_preds)

#Model2tx MSPE
MSPETX2 = mean((TestTX$ABV - model2tx_preds)^2)
MSPETX2
```

##### Check stability of Model 1 vs Model 2 Texas
```{r}
numMSPEs = 1000
MSPEHolderModel1 = numeric(numMSPEs)
MSPEHolderModel2 = numeric(numMSPEs)

for (i in 1:numMSPEs)
{
  TrainObs = sample(seq(1,dim(beerTX)[1]),round(.60*dim(beerTX)[1]), replace = FALSE)
  TrainingTX <- beerTX[TrainObs,]
  TrainingTX
  TestTX <- beerTX[-TrainObs,]
  TestTX
  model1tx <- lm(ABV ~ IBU, data = TrainingTX)
  model1tx_preds <- predict(model1tx, newdata = TestTX)
  
  #MSPE Model 1
  MSPETX1 = mean((TestTX$ABV - model1tx_preds)^2)
 MSPETX1
  MSPEHolderModel1[i] = MSPETX1
  
  #Model 2
  model2tx <- lm(ABV ~ IBU + IBU2, data = TrainingTX)
  model2tx_preds <- predict(model2tx, newdata = TestTX)
  MSPETX2 = mean((TestTX$ABV - model2tx_preds)^2)
  MSPETX2
  MSPEHolderModel2[i] = MSPETX2
  
}

mean(MSPEHolderModel1)
mean(MSPEHolderModel2)

```

##### Model 1 Colorado
```{r}
#Train model1 Texas
model1co <- lm(ABV ~ IBU, data = TrainingCO)

#Predictions based on model1tx
model1co_preds <- predict(model1co, newdata = TestCO)
as.data.frame(model1co_preds)

#Model1tx MSPE
MSPECO1 = mean((TestCO$ABV - model1co_preds)^2)
MSPECO1
```

##### Model 2 Colorado
```{r}
#Train model2 Texas
model2co <- lm(ABV ~ IBU + IBU2, data = TrainingCO)

#Predictions based on model1tx
model2co_preds <- predict(model2co, newdata = TestCO)
as.data.frame(model2co_preds)

#Model2tx MSPE
MSPECO2 = mean((TestCO$ABV - model2co_preds)^2)
MSPECO2
```


##### Check stability of Model 1 vs Model 2 Colorado
```{r}
numMSPEs = 1000
MSPEHolderModel1 = numeric(numMSPEs)
MSPEHolderModel2 = numeric(numMSPEs)

for (i in 1:numMSPEs)
{
  TrainObs = sample(seq(1,dim(beerCO)[1]),round(.60*dim(beerCO)[1]), replace = FALSE)
  TrainingCO <- beerCO[TrainObs,]
  TrainingCO
  TestCO <- beerCO[-TrainObs,]
  TestCO
  model1co <- lm(ABV ~ IBU, data = TrainingCO)
  model1co_preds <- predict(model1co, newdata = TestCO)
  
  #MSPE Model 1
  MSPECO1 = mean((TestCO$ABV - model1co_preds)^2)
 MSPECO1
  MSPEHolderModel1[i] = MSPECO1
  
  #Model 2
  model2co <- lm(ABV ~ IBU + IBU2, data = TrainingCO)
  model2tx_preds <- predict(model2co, newdata = TestCO)
  MSPECO2 = mean((TestCO$ABV - model2co_preds)^2)
  MSPECO2
  MSPEHolderModel2[i] = MSPECO2
  
}

mean(MSPEHolderModel1)
mean(MSPEHolderModel2)

```
```

##### ASE loss function I'm note sure how this helps me decide which model is better? Please explaine. I get that running the mspe in a thosand draws can show stability of a model by how much the value fluctuates. 