---
title: "Homework_Unit5"
author: "Matt Chinchilla"
date: "6/7/2019"
output: html_document
Keep_md: True
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1a

```{r}
#Download the yob2016.txt data
path <- "https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%205/yob2016.txt"
download.file(path, destfile = "yob2016.txt")
df <- read.delim("yob2016.txt", header = FALSE, sep = ";")

#View head data
head(df)

#Rename column headers to human readable
names(df)[1] <- "First Name"
names(df)[2] <- "Gender"
names(df)[3] <-"Amount of childern"
head(df)
```

# 1b
```{r}
#Summary of df
summary(df)

#Sturcture of df
str(df)
```

# 1c find name that ends in three y's
```{r}
#Find a patern of three y's in the First Name column
grep("yyy", df$`First Name`)

#View the row with the misspelled name
df[212,]
```

# 1d remove the misspelled name
```{r}
#Install the dplyr library to use slice 
library(dplyr)
#remove the row with the misspelled word and recalculate ".preserve = FALSE" the position values after removing the rows. Name this new daata frame y2016
y2016 <- slice(df, -c(212), .preserve = FALSE)

#check to make sure the row of data has been removed by running the patern matching again
grep("yyy", y2016$`First Name`)
```

# 2a Import yob2015.txt data
```{r}
#URL of yob2015.txt data
path <- "https://raw.githubusercontent.com/BivinSadler/MSDS-6306-Doing-Data-Science/master/Unit%205/yob2015.txt"

#Download yob2015.txt to working directory
download.file(path, destfile = "yob2015.txt")

#Create a data frame from the yob2015 data called y2015
y2015 <- read.delim("yob2015.txt", header = FALSE, sep = ",")

#Make the column names in y2015 data frame human readable
names(y2015)[1] <- "First Name"
names(y2015)[2] <- "Gender"
names(y2015)[3] <-"Amount of childern"
```

# 2b Display the last ten rows of the y2015 data frame
#### Observation of last ten rows: The last ten rows are all males and there are only five children with each name.
```{r}
tail(y2015, 10)
```

# 2c Merge y2015 and y2016 by first name
```{r}
#Create a new data frame called final. Inner join y2015 and y2016 data frames on First name
final <- merge(y2015, y2016, by = "First Name")

#View the first few rows of final
head(final)

#Remove the extra row containing gender info
final <- subset(final, select = -c(Gender.y))

#View the first few rows of final
head(final)

#rename the last three columns "Gender", "Amount of Children 2015", "Amount of Children 2016"
names(final)[2] <- "Gender"
names(final)[3] <-"Amount of childern 2015"
names(final)[4] <-"Amount of childern 2016"

#View the first few rows of final
head(final)
```

# 3a Create a total column
```{r}
#Create a new column called total with the number of people given each name combined for both years.
final$Total <- final[,3] + final[,4]

#See the first few rows.
head(final)
```

# 3b Sort the data and display the top 10 names
```{r}
#Create a new data frame called Finaldec that shows names in order of most popular to least. 
Finaldec <- final[order(-final$Total),]

#Show the top ten names
head(Finaldec, 10)
```

# 3c Top 10 girl names
```{r}
#Create a new data set called JustGirls that filters boys out of the finaldec data set.
JustGirls <- filter(Finaldec, Gender == "F")

#View the top 10 girl names.
head(JustGirls, 10)
```

# 3c Create top 10 girl names CSV
```{r}
#Create an new Data frame called JustGirlsTop10 that has the top ten girl names
JustGirlsTop10 <- JustGirls[c(1:10),]

#Subset the data to only include First name and Total
JustGirlsTop10 <- subset(JustGirlsTop10, select = -c(Gender, `Amount of childern 2015`, `Amount of childern 2016`))
JustGirlsTop10

#Write CSV file called GirlsTop10Names
write.csv(JustGirlsTop10, file = "GirlsTop10Names.csv")
```
