---
title: "Homework_Unit6"
author: "Matt Chinchilla"
date: "6/14/2019"
output: html_document
Keep_md: True
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1b State abbreviations without their counts.
```{r}
#Load the data
load(file = "./N-MHSS-2015-DS0001-data-r.rda")

#Knitr library
library(knitr)

#Find unique state abbreviations (Factors)
st <- unique(mh2015_puf$LST)

#Display list of unique state abbreviations (Display factors in html friendly format)
kable(head(st), format = "html", col.names = "States")
```

### 1c Veterans Veterans Administration (VA) medical centers in the mainland United States

```{r}
#dplyr Library
library(dplyr)

#Create a subset of mh2015 data to isolate only the VA facilitiesn
VA <- subset(mh2015_puf, FACILITYTYPE == "Veterans Administration medical center (VAMC) or other VA health care facility")

#Remove all rows with states or territories outside of the US mainland
VA2 <- VA %>% filter(!LST %in% c("HI    ", "AK    ", "VI    ", "PR    " ))

#Create a data frame of counts of VA hospitals in each state called VACount
VACount <- VA2 %>% count(LST)

#Rename the columns VACount to State and Count
names(VACount) <- c("State", "Count")

kable(head(VACount), format = "html")
```

### 1d Create a ggplot barchart of filtered data
```{r}
#Load ggplot2 library
library(ggplot2)

#Create Bar chart of VACount data
VABar <- ggplot(VACount, aes(reorder(VACount$State, -VACount$Count), y = VACount$Count, fill = VACount$State)) + geom_bar(stat = "identity") +ggtitle("VA Count Per State")+ theme(legend.position = "none", plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 90)) + labs(x = "State", y = "Count of VA Hospitals") 

#Display Bar chart
VABar
```

### 2a Observation: There are extra spaces in the formating of the state abbreviations in the VA data.
```{r}
#Download statesize.csv file from https://raw.githubusercontent.com/Chinchillin1981/MSDS-6306-Doing-Data-Science/master/Unit%206/statesize.csv
#download.file(URL, destfile = "statesize.csv")

#Read statesize.csv file into global environment
statesize <- read.csv(file = "./statesize.csv")

#Look at format of State in VACount 
head(paste(VACount$State))

#Look at format of State in statesize
head( paste(statesize$Abbrev))
```

### 2b 
```{r}
#Trim the white spaces on the State abbreviations in VACount
VACount[] <- lapply(VACount, function(x) if (is.factor(x)) sub(" +$", "",x) else x)

#Check trim
head(paste(VACount$State))
```

```{r}
#Merge VACount and Statesize 
VAState <- merge(VACount,statesize, by.x = "State", by.y = "Abbrev", all.x = TRUE)

#Check merge data
head(VAState)
```


### 2c Create a new column that is a calculation of the number of VA facilites per 1000 square miles in each state.
```{r}
#Create a new column VAPerThousandSqMiles 
VAState$VAPerThousandSqMiles <- VAState$Count/(VAState$SqMiles/1000)

head(VAState)
```

### 2d Plot VAs per 1000 square mile in a bar plot.
```{r}
#Load ggplot Library
library(ggplot2)

#Plot the VA Hospitals Per 1000 sqMile
p<-ggplot(VAState ,aes(reorder(VAState$State,-VAPerThousandSqMiles),VAPerThousandSqMiles, fill=Region)) + geom_bar(stat='identity')+ theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 90))+ ggtitle("VA Per 1000 SqMiles in Each State") + labs(x = "State", y = "Number of VA Per 1000 SqMile") + scale_fill_manual(values = c("Blue", "red","orange","green"))

#Display Plot
p
```

### 2e 
##### Observations:
+ The Northeast region has the highest number of VA's per 1000 SqMiles 
+ The West region has a very low number of VA's per 1000 SqMiles

##### Advice:
My advice to my boss at this point would be that we should gather information related to Veterans and where they are located. The thinking being that we should take the information we have gathered so far and focus on finding areas that have a high concentration of Veterans and no VA center within at least a 1000 mile radius. These areas are probably the locations where a new VA hospital would do the most good. 

